---
title: "p8105_hw2_yd2739"
author: "Yuxuan Du"
date: "2023-09-26"
output: github_document
---

## Problem 1

We'll use the data in `pols-month.csv`, `unemployment.csv`, and `snp.csv`

**load `pols-month.csv`**
```{r}
library(tidyverse)
pols_month_df = 
  read_csv("HW2_P1_DATA/pols-month.csv")|>
  janitor::clean_names()|>
  separate(
    col = mon,
    into = c("year", "month", "day"), 
    sep = "-") |>
  mutate(
    month = case_match(
      month, 
      '01' ~ "January", 
      '02' ~ "February", 
      '03' ~ "March", 
      '04' ~ "April", 
      '05' ~ "May", 
      '06' ~ "June", 
      '07' ~ "July", 
      '08' ~ "August", 
      '09' ~ "September", 
      '10' ~ "October", 
      '11' ~ "November", 
      '12' ~ "December"
    )
  )|>
  mutate(
    president = case_match(
      prez_gop, 
      2 ~ "republican",
      1 ~ "republican", 
      0 ~ "democratic"
    )
  )|>
  subset(select = -c(prez_gop, prez_dem, day))
pols_month_df
```

**load `snp.csv`**
```{r}
snp_df = 
  read_csv("HW2_P1_DATA/snp.csv")|>
  janitor::clean_names()|>
  separate(
    col = date,
    into = c("month", "day", "year"), 
    sep = "/") |>
  mutate(
    month = case_match(
      month, 
      '1' ~ "January", 
      '2' ~ "February", 
      '3' ~ "March", 
      '4' ~ "April", 
      '5' ~ "May", 
      '6' ~ "June", 
      '7' ~ "July", 
      '8' ~ "August", 
      '9' ~ "September", 
      '10' ~ "October", 
      '11' ~ "November", 
      '12' ~ "December"
    )
  )|>
  mutate(
    year = as.numeric(year)
  )|>
  mutate(
    year = ifelse(year>=50, 1900+year, 2000+year), 
    year = as.character(year)
  )|>
  relocate(c(year, month))
snp_df
```

**load`unemployment.csv`**
```{r}
unemployment_df = 
  read_csv("HW2_P1_DATA/unemployment.csv")|>
  janitor::clean_names()|>
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment", 
  )|>
  mutate(
    year = as.character(year)
  )|>
  mutate(
    month = case_match(
      month, 
      'jan' ~ "January", 
      'feb' ~ "February", 
      'mar' ~ "March", 
      'apr' ~ "April", 
      'may' ~ "May", 
      'jun' ~ "June", 
      'jul' ~ "July", 
      'aug' ~ "August", 
      'sep' ~ "September", 
      'oct' ~ "October", 
      'nov' ~ "November", 
      'dec' ~ "December"
    )
  )
unemployment_df
```
**merge three dataframe**
```{r}
P1_whole_df = 
  left_join(pols_month_df, snp_df, by = c("year", "month"))|>
  left_join(unemployment_df, by = c("year", "month"))
P1_whole_df
```

So the pols_month dataset contains the information of president, governors, senators, and representatives from republican and democratic during different period of time. 

The snp dataset contains the Standard & Poor’s stock market index during different period of time. 

The unemployment dataset contains the unemployment rate in the US during different period of time. 

The time span of the dataset is from 1947 to 2015. The size of the merged dataset is 822 x 12. 



## Problem 2
```{r}
# load the dataset, change the format of date, recalculate homes_powered
mr_trash_wheel_df = 
  readxl::read_excel("HW2_P2_DATA/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N586")|>
  janitor::clean_names()|>
  mutate(
    date = substr(date, nchar(date) - 1, nchar(date))
  )|>
  mutate(
    homes_powered = (weight_tons*500/30)
  )|>
  mutate(name = "Mr_trash_wheel")
mr_trash_wheel_df
```
```{r}
professor_trash_wheel_df = 
  readxl::read_excel("HW2_P2_DATA/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M108")|>
  janitor::clean_names()|>
  mutate(
    date = substr(date, nchar(date) - 1, nchar(date))
  )|>
  mutate(
    homes_powered = (weight_tons*500/30), 
    year = as.character(year)
  )|>
  mutate(name = "Professor_trash_wheel")
professor_trash_wheel_df
```

```{r}
gwynnda_trash_wheel_df = 
  readxl::read_excel("HW2_P2_DATA/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L157")|>
  janitor::clean_names()|>
  mutate(
    date = substr(date, nchar(date) - 1, nchar(date))
  )|>
  mutate(
    homes_powered = (weight_tons*500/30),
    year = as.character(year)
  )|>
  mutate(name = "Gwynnda_trash_wheel")

trash_wheel_collection_df = 
  bind_rows(mr_trash_wheel_df, professor_trash_wheel_df, gwynnda_trash_wheel_df)

total_weight_professor = sum(pull(professor_trash_wheel_df, weight_tons))

total_cigarette_butts_gwynnda_july_2021 = subset(
  gwynnda_trash_wheel_df, 
  month == "July" & year == "2021"
  )|>
  pull(
    cigarette_butts
    )|>
  sum()
gwynnda_trash_wheel_df
```

The size of the merged dataframe is 845 x 15.  

Some interesting information: Only Mr.Trash Wheel has the variable of `sports_ball`;

Gwynnda Trash Wheel does not have the variable `galss_bottles`;

The way of recording date is kind of redundant since there are variables for year and month (the format of date is changed later)

There might be missing data in the dataset gwynnda_trash_wheel_df since there are days they collected tons of trash while all trash category counts is 0. 

The total weight of trash collected by Professor Trash Wheel is `r total_weight_professor`. 

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r total_cigarette_butts_gwynnda_july_2021`. 

## Problem 3

```{r}
demo_info_df = 
  read_csv("HW2_P3_DATA/MCI_baseline.csv", skip = 1)|>
  janitor::clean_names()|>
  mutate(
    sex = recode(sex, "1" = "male", "0" = "female"),
    apoe4 = recode(apoe4, "1" = "carrier", "0" = "non-carrier"), 
    sex = factor(sex), 
    apoe4 = factor(apoe4)
  )|>
  subset(
    current_age < age_at_onset | age_at_onset == "."
  )|>
  rename(
    age_starting = current_age
  )|>
  mutate(
    age_at_onset = as.numeric(ifelse(age_at_onset ==".", NA, age_at_onset))
  )
head(demo_info_df)

got_MCI_df = demo_info_df |>
  subset(
    age_at_onset != "Na"
  )

avg_age_baseline = mean(pull(demo_info_df, age_starting))

total_women = subset(
  demo_info_df, 
  sex == "female"
  )|>
  nrow()

total_women_carrier = subset(
  demo_info_df, 
  sex == "female" & apoe4 == "carrier"
  )|>
  nrow()

women_apoe4_percent = total_women_carrier/total_women
```

We can see that there are `r nrow(read_csv("HW2_P3_DATA/MCI_baseline.csv", skip = 1))` people recruited in the research. Among these people, there are `r nrow(demo_info_df)` people met the inclusion criteria that didn't have MCI at baseline and joined the experiment. Among the 479 participants, `r nrow(got_MCI_df)` of the `r nrow(demo_info_df)` developed MCI at last. The average baseline age is `r avg_age_baseline`. The proportion of women in the study who are APOE4 carriers is `r women_apoe4_percent`. There are lot of participant who didn't got MCI at the end of the research, and their onset age is recorded as '.', these values were replaced by Na, but I didn't filter them out since it might still be useful if we want to do analysis using models like logistic regression to analyze the relationship of different variables and their final MCI status. To filter them out, we only need to use drop_na() function.

```{r}
amyloid_df = 
  read_csv("HW2_P3_DATA/mci_amyloid.csv", skip = 1)|>
  janitor::clean_names()|>
  rename(
    id=study_id, 
    baseline = baseline, 
    second_year = time_2, 
    fourth_year = time_4, 
    sixth_year = time_6, 
    eighth_year = time_8)

amyloid_pivoted_df = amyloid_df |>
  pivot_longer(
    baseline:eighth_year, 
    names_to = "time",
    values_to = "42/40_ratio", 
  )
head(amyloid_df)
head(amyloid_pivoted_df)
```

We also need to skip the first line of mci_amyloid.csv since it is not the header for dataframe. Also to follow the format of `demo_info_df`, the variable `study_id` is renamed as `id`. There are missing values in dataset, it might means those participants didn't show up at that time. The size of the dataframe is 487 x 6. Also, we could pivot the dataset longer, it was named as `amyloid_pivoted_df`. There are `r nrow(amyloid_df)` observations in `amyloid_df`, if we pivot it, there are `r nrow(amyloid_pivoted_df)` observations in  `amyloid_pivoted_df`. (since according to discussion, there are no inclusion criteria for amyloid dataset)

```{r}
only_baseline = demo_info_df |>
  anti_join(amyloid_df, by = "id")
head(only_baseline)

only_amyloid = amyloid_df |>
  anti_join(demo_info_df, by = "id")
head(only_amyloid)

both_df = merge(demo_info_df, amyloid_df)

both_pivoted_df = merge(demo_info_df, amyloid_pivoted_df)

head(both_df)

head(both_pivoted_df)

write.csv(both_df, "HW2_P3_DATA/merged.csv", row.names = FALSE)

write.csv(both_pivoted_df, "HW2_P3_DATA/merged_pivoted.csv", row.names = FALSE)

female_count = nrow(filter(both_df, sex == "female"))

male_count = nrow(filter(both_df, sex == "male"))

MCI_count = nrow(filter(both_df, age_at_onset != '.'))
```

There are 16 participants appear only in the amyloid dataset and 8 participants appear only in the baseline dataset. 

For the merged dataset, there are `r nrow(both_df)` observations. 

For the pivoted merged dataset, there are `r nrow(both_pivoted_df)` observations.(since each participant have 5 42/40 beta amyloid ratio taken at different times)

There are `r female_count` female participant and `r male_count` male participant both appeared in two dataset, and there are `r MCI_count` participant who developed MCI at last. 