---
title: "p8105_hw2_yd2739"
author: "Yuxuan Du"
date: "2023-09-26"
output: github_document
---

## Problem 1

We'll use the data in `pols-month.csv`, `unemployment.csv`, and `snp.csv`

**load `pols-month.csv`**
```{r}
library(tidyverse)
pols_month_df = 
  read_csv("HW2_P1_DATA/pols-month.csv")|>
  janitor::clean_names()|>
  separate(
    col = mon,
    into = c("year", "month", "day"), 
    sep = "-") |>
  mutate(
    month = case_match(
      month, 
      '01' ~ "January", 
      '02' ~ "February", 
      '03' ~ "March", 
      '04' ~ "April", 
      '05' ~ "May", 
      '06' ~ "June", 
      '07' ~ "July", 
      '08' ~ "August", 
      '09' ~ "September", 
      '10' ~ "October", 
      '11' ~ "November", 
      '12' ~ "December"
    )
  )|>
  mutate(
    president = case_match(
      prez_gop, 
      1 ~ "republican", 
      0 ~ "democratic"
    )
  )|>
  subset(select = -c(prez_gop, prez_dem, day))
```

**load `snp.csv`**
```{r}
snp_df = 
  read_csv("HW2_P1_DATA/snp.csv")|>
  janitor::clean_names()|>
  separate(
    col = date,
    into = c("month", "day", "year"), 
    sep = "/") |>
  mutate(
    month = case_match(
      month, 
      '1' ~ "January", 
      '2' ~ "February", 
      '3' ~ "March", 
      '4' ~ "April", 
      '5' ~ "May", 
      '6' ~ "June", 
      '7' ~ "July", 
      '8' ~ "August", 
      '9' ~ "September", 
      '10' ~ "October", 
      '11' ~ "November", 
      '12' ~ "December"
    )
  )|>
  mutate(
    year = as.numeric(year)
  )|>
  mutate(
    year = ifelse(year>=50, 1900+year, 2000+year), 
    year = as.character(year)
  )|>
  relocate(c(year, month))
```

**load`unemployment.csv`**
```{r}
unemployment_df = 
  read_csv("HW2_P1_DATA/unemployment.csv")|>
  janitor::clean_names()|>
  pivot_longer(
    jan:dec, 
    names_to = "month",
    values_to = "unemployment", 
  )|>
  mutate(
    year = as.character(year)
  )|>
  mutate(
    month = case_match(
      month, 
      'jan' ~ "January", 
      'feb' ~ "February", 
      'mar' ~ "March", 
      'apr' ~ "April", 
      'may' ~ "May", 
      'jun' ~ "June", 
      'jul' ~ "July", 
      'aug' ~ "August", 
      'sep' ~ "September", 
      'oct' ~ "October", 
      'nov' ~ "November", 
      'dec' ~ "December"
    )
  )
  
```
**merge three dataframe**
```{r}
P1_whole_df = 
  left_join(pols_month_df, snp_df, by = c("year", "month"))|>
  left_join(unemployment_df, by = c("year", "month"))
```

So the pols_month dataset contains the information of president, governors, senators, and representatives from republican and democratic during different period of time. 

The snp dataset contains the Standard & Poor’s stock market index during different period of time. 

The unemployment dataset contains the unemployment rate in the US during different period of time. 

The time span of the dataset is from 1947 to 2015. The size of the merged dataset is 822 x 12. 
